---
title: "enron_exploration"
author: "Mark Ayzenshtadt"
date: "30 Dec 2015"
output: html_document
---

```{r}
setwd('~/DataAnalyst/Projects/DataAnalystND_Project_5/ud120-projects/final_project/')
enron = read.csv('final_project_data.csv')
enron <- subset(enron, name != 'TOTAL')
enron <- enron[-18]
subset(enron, is.nan) <- 0
library(ggplot2)
```


```{r, echo=FALSE}
ggplot(enron, aes(salary, from_poi_to_this_person, color = poi)) +
  geom_point()
```

```{r}
feature_stats <- data.frame(row.names = names(enron))
for (col in names(enron)) {
  feature_stats[col, 'notNaN'] <- nrow(subset(enron, enron[[col]] != 'NaN'))
  feature_stats[col, 'unique_values'] <- length(unique(enron[[col]]))
}
```

Def-payment - deferreed income
Total payments - other
total stock value - excercised stock options
total stock value - restricted stock
I'll retain only total_stock_value as it has less NaNs
to_messages - shared reciept with poi
```{r}
enron1 <- enron
enron1$poi <- as.integer(enron1$poi) - 1
View(cor(enron1[c(-1)], use = 'pairwise.complete.obs'))
```

Remove:
deferral_payments - corr with deferred_income
loan_advances - only 3 values
excercised_stock_options - contained in total_stock_value
other - corr with total_payments
restricted_stock - contained in total_stock_value
shared_receipt_with_poi - corr with to_messages
```{r}
names(enron)
enron <- enron[c(-4,-6,-12,-13,-15,-22)]
```

I expect money not to be paid if financial data is NA.
Same for emails - NA -> no emails present.
```{r}
enron[is.na(enron)] <- 0
enron1 <- enron
enron1$poi <- as.integer(enron1$poi) - 1
View(cor(enron1[c(-1)], use = 'pairwise.complete.obs'))
```

18 are poi, 127 not
```{r}
summary(enron$poi)
```

Few outliers in salary, others clustered.
```{r}
ggplot(enron, aes(salary)) +
  geom_histogram()
```

1 outlier - Ken Lay, most have very low total_payments.
```{r}
summary(enron$total_payments)
ggplot(enron, aes(total_payments)) +
  geom_histogram()

ggplot(subset(enron, total_payments < 9e+07), aes(total_payments)) +
  geom_histogram()
subset(enron, total_payments > 9e+07)$name
```

Outliers - Lavorato, Lay, Belden, Skilling
Most < 2mil.
```{r}
summary(enron$bonus)
ggplot(enron, aes(bonus)) +
  geom_histogram(binwidth = 100000)
subset(enron, bonus > 5e6)$name
```

Outliers - Hannon, Belden, Rice, Allen, Frevert
```{r}
summary(enron$deferred_income)
ggplot(enron, aes(deferred_income)) +
  geom_histogram()
ggplot(enron, aes(poi, deferred_income)) +
  geom_boxplot() +
  stat_summary(fun.y = mean, geom = 'point', shape = 5)

subset(enron, deferred_income < -2e6)$name
```

Outliers - Ken, Rice, Skilling, Hirko, Pai
Most < 2.5mil
```{r}
summary(enron$total_stock_value)
ggplot(enron, aes(total_stock_value)) +
  geom_histogram()
ggplot(enron, aes(total_stock_value)) +
  geom_histogram()
subset(enron, total_stock_value > 2e7)$name
```

Expenses - ok
```{r}
summary(enron$expenses)
ggplot(enron, aes(expenses)) +
  geom_histogram()
```

Outliers - Lay, Martin
```{r}
summary(enron$long_term_incentive)
ggplot(enron, aes(long_term_incentive)) +
  geom_histogram()

subset(enron, long_term_incentive > 3e6)$name
```

Outliers - Shapiro, Kean
```{r}
summary(enron$to_messages)
ggplot(enron, aes(to_messages)) +
  geom_histogram()

subset(enron, to_messages > 1e4)$name
```

Outlier - Lavorato
```{r}
summary(enron$from_poi_to_this_person)
ggplot(enron, aes(from_poi_to_this_person)) +
  geom_histogram()

subset(enron, from_poi_to_this_person > 400)$name
```

Outliers - Beck, Kaminsky, Kean
```{r}
summary(enron$from_messages)
ggplot(enron, aes(from_messages)) +
  geom_histogram()

subset(enron, from_messages > 4000)$name
```

Outliers - Beck, Lavorato, Delainey, Kean
```{r}
summary(enron$from_this_person_to_poi)
ggplot(enron, aes(from_this_person_to_poi)) +
  geom_histogram()

subset(enron, from_this_person_to_poi > 300)$name
```


```{r}
summary(enron$deferral_payments)
ggplot(enron, aes(deferral_payments)) +
  geom_histogram()

subset(enron, deferral_payments > 2e6)$name
```


```{r}
summary(enron$restricted_stock_deferred)
ggplot(enron, aes(restricted_stock_deferred)) +
  geom_histogram()

subset(enron, restricted_stock_deferred > 1e7)$name
```

```{r}
summary(enron$exercised_stock_options)
ggplot(enron, aes(exercised_stock_options)) +
  geom_histogram()

subset(enron, exercised_stock_options > 1e7)$name
```

```{r}
summary(enron$other)
ggplot(enron, aes(other)) +
  geom_histogram()

subset(enron, other > 1e7)$name
```

